---
title: "Analysis of ctDNA Data from GIST  Phase I Study"
author: "Stat4ward, `r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document:
     reference_docx: reference_doc.docx
toc: TRUE
---
\pagebreak

# GOAL

# DATA

```{r setup,include=FALSE}
knitr::opts_chunk$set(cache=FALSE,echo = FALSE, warning=FALSE, message=FALSE, fig.width = 6.5, fig.height = 6.5)

rm(list = ls())
library(haven)
library(tidyverse)
library(readr)
library(ggplot2)
library(knitr)
library(readxl)
library(purrr)
library(tableone)

# data input

adrs<- read_sas("C:\\Users\\ShaYang\\OneDrive - Stat4ward\\Documents\\Deciphera\\Phase I\\GIST\\adrs.sas7bdat")
#glimpse(adrs)

adsl<- read_sas("C:\\Users\\ShaYang\\OneDrive - Stat4ward\\Documents\\Deciphera\\Phase I\\GIST\\adsl.sas7bdat")
#glimpse(adsl)

adtte<-read_sas("C:\\Users\\ShaYang\\OneDrive - Stat4ward\\Documents\\Deciphera\\Phase I\\GIST\\adtte.sas7bdat")
#glimpse(adtte)

ctDNA.all<-read.csv("C:\\Users\\ShaYang\\OneDrive - Stat4ward\\Documents\\Deciphera\\Phase I\\GIST\\20210209_Deciphera_SOW1_1064Samples_G360Report_Cumulative.csv")

ctDNA<-ctDNA.all
# %>%filter(Cancertype=="GIST (Gastrointestinal stromal tumor)"|Cancertype=="GIST (Gastrointestinal Stromal Tumor)")

```

ctDNA data from Guardant Health (20201223_Deciphera_SOW1_1064Samples_G360Report_Cumulative.csv)
```{r}

ctDNA<-ctDNA%>%mutate(Patient_ID=gsub(" ","", Patient_ID))

no.patient<-n_distinct(ctDNA$Patient_ID)

ctDNA1<-ctDNA%>%
  filter(Sample_status=="SUCCESS")

no.row.fail<-nrow(ctDNA)-nrow(ctDNA1)

ctDNA.germline<-nrow(ctDNA1%>%filter(Somatic_status=="germline"))

# ctDNA1<-ctDNA1%>%
#   filter(Somatic_status=="somatic")  will update this once the somatic status is corrected

 ctDNA1<-ctDNA1%>%
   filter(Somatic_status!="germline")  

   
#no.patient.somatic<-n_distinct(ctDNA1$Patient_ID)

#unique(ctDNA1$Variant_type)

#ctDNA1%>%filter(Variant_type=="Fusion")

#ctDNA.fusion<-ctDNA1%>%filter(Variant_type=="Fusion")

#ctDNA1<-ctDNA1%>%filter(Variant_type!="Fusion")%>%droplevels()

no.patient.sucess<-n_distinct(ctDNA1$Patient_ID)

ctDNA1<-ctDNA1%>%mutate(Visit_name=gsub(" ","",Visit_name))%>%
  mutate(Visit_name=gsub("BASELINE", "C1D1", Visit_name,ignore.case = TRUE))%>%
  mutate(Visit_name=gsub("UNSCH","Unscheduled",Visit_name))%>%
  mutate(Visit_name=gsub("UNS","Unscheduled",Visit_name))%>%
  mutate(Visit_name=gsub("Day1","D1", Visit_name))

ctDNA1<-ctDNA1%>%mutate(Received_date=as.Date(Received_date, "%m/%d/%Y"),
                   Bloodcoll_date=as.Date(Bloodcoll_date, "%m/%d/%Y"),
                   Reported_date=as.Date(Reported_date,"%m/%d/%Y"))

ctDNA1<-ctDNA1%>%mutate(Cancer.type=case_when(
  Cancertype=="GIST (Gastrointestinal Stromal Tumor)"~ "GIST",
  Cancertype=="GIST (Gastrointestinal stromal tumor)"~ "GIST",
  Cancertype=="Glioblastoma"~ "GBM",
  TRUE ~ Cancertype))

ctDNA1<-ctDNA1%>%mutate(Cancer.type = na_if(Cancer.type, ""))
```

There are `r nrow(ctDNA)` rows in the original ctDNA data set for Cancertype GIST (Gastrointestinal stromal tumor). `r no.row.fail` rows are excluded since Sample Status is NEGATIVE, `r n_distinct(ctDNA$Patient_ID)-n_distinct(ctDNA1$Patient_ID)` patients are eliminated because of this step. There are  `r no.patient.sucess` patients in the data with Sample Status being Success. 

```{r}

visit.summary<-as.data.frame.matrix(table(ctDNA1$Patient_ID, ctDNA1$Visit_name))

myVars = c("Variant_type","Indel_type","Gene","Exon","Percentage","Received_date")
catVars= c("Variant_type","Indel_type","Gene","Exon","Received_date")
tab.summary <- CreateTableOne(vars = myVars, data = ctDNA1, factorVars = catVars)

```


```{r}

#Imputations

ctDNA1<-ctDNA1%>%group_by(Patient_ID)%>%fill(Cancer.type, .direction = "updown")%>%ungroup()

ctDNA.mutation<-ctDNA1%>%
  filter(Variant_type!="CNV")%>%
  #mutate(Visit_name=droplevels(as.factor(Visit_name)), Gene=droplevels(as.factor(Gene)), Exon=droplevels(as.factor(Exon)))%>%
  expand(Visit_name,nesting(Patient_ID,Variant_type,Gene, Exon, Mut_aa, Mut_nt))%>%arrange(Patient_ID,Visit_name,Variant_type,Gene,Exon)

visit.all<-paste(ctDNA1$Patient_ID, ctDNA1$Visit_name)
visit.check<-unique(visit.all)

impute<-ctDNA.mutation%>%left_join(ctDNA1, by = c("Patient_ID"="Patient_ID",
                                                        "Visit_name"="Visit_name",
                                                        "Variant_type"="Variant_type",
                                                        "Gene"="Gene",
                                                        "Exon"="Exon",
                                                        "Mut_aa"="Mut_aa",
                                                        "Mut_nt"="Mut_nt"))%>%
  group_by(Patient_ID,Variant_type,Gene,Exon,Mut_aa,Mut_nt)%>%
  fill(`ï..Study_ID`:Transcript, .direction = "updown")%>%
  fill(Cancertype, .direction = "updown")%>%
  ungroup()%>%
  mutate(Percentage=case_when(
    !is.na(Percentage)~Percentage,
    is.na(Percentage)&paste(Patient_ID,Visit_name)%in%visit.check ~ 0.02,
    is.na(Percentage)&!paste(Patient_ID,Visit_name)%in%visit.check ~ 999))%>%
  filter(Percentage!=999)

# %>%mutate(Visit_name=
#         factor(Visit_name,levels =       c("C1D1","C3D1","C5D1","C5D15","C7D1","C9D1","C11D1","C12D1","COC1D1","COC3D1","COC5D1","COC7D1","COC9D1","EOT")) )


```