---
title: "Analysis of ctDNA Data from GIST  Phase I Study"
author: "Stat4ward, `r format(Sys.time(), '%d %B, %Y')`"
output:
  word_document:
     reference_docx: reference_doc.docx
toc: TRUE
---
\pagebreak

# GOAL

# DATA

```{r setup,include=FALSE}
knitr::opts_chunk$set(cache=FALSE,echo = FALSE, warning=FALSE, message=FALSE, fig.width = 6.5, fig.height = 6.5)

rm(list = ls())
library(haven)
library(tidyverse)
library(readr)
library(ggplot2)
library(knitr)
library(readxl)
library(purrr)

# data input

adrs<- read_sas("C:\\Users\\Sha\\OneDrive\\Documents\\Deciphera\\Phase I Study\\GIST\\adrs.sas7bdat")
#glimpse(adrs)

adsl<- read_sas("C:\\Users\\Sha\\OneDrive\\Documents\\Deciphera\\Phase I Study\\GIST\\adsl.sas7bdat")
#glimpse(adsl)

adtte<-read_sas("C:\\Users\\Sha\\OneDrive\\Documents\\Deciphera\\Phase I Study\\GIST\\adtte.sas7bdat")
#glimpse(adtte)

ctDNA.all<-read.csv("C:\\Users\\Sha\\OneDrive\\Documents\\Deciphera\\Phase I Study\\GIST\\20201223_Deciphera_SOW1_1064Samples_G360Report_Cumulative.csv")

ctDNA<-ctDNA.all%>%
  filter(Cancertype=="GIST (Gastrointestinal stromal tumor)"|Cancertype=="GIST (Gastrointestinal Stromal Tumor)")

```

ctDNA data from Guardant Health (20201223_Deciphera_SOW1_1064Samples_G360Report_Cumulative.csv)
```{r}

no.patient<-n_distinct(ctDNA$Patient_ID)

ctDNA1<-ctDNA%>%
  filter(Sample_status=="SUCCESS")

no.row.fail<-nrow(ctDNA)-nrow(ctDNA1)

#ctDNA.germline<-ctDNA1%>%filter(Somatic_status=="germline")

# no.row.gerline<-nrow(ctDNA1)-nrow(ctDNA1<-ctDNA1%>%
#   filter(Somatic_status=="somatic"))
#   
# no.patient.somatic<-n_distinct(ctDNA1$Patient_ID)

#unique(ctDNA1$Variant_type)

#ctDNA1%>%filter(Variant_type=="Fusion")

#ctDNA.fusion<-ctDNA1%>%filter(Variant_type=="Fusion")

ctDNA1<-ctDNA1%>%filter(Variant_type!="Fusion")%>%droplevels()

no.patient.sucess<-n_distinct(ctDNA1$Patient_ID)


ctDNA1%>%filter(Gene=="PDGFRA")%>%select(Patient_ID, Visit_name,Gene, Exon )%>%view()

summary(ctDNA1$Percentage)

```

There are `r nrow(ctDNA)` rows in the original ctDNA data set for Cancertype GIST (Gastrointestinal stromal tumor). `r no.row.fail` rows are excluded since Sample Status is NEGATIVE, `r n_distinct(ctDNA$Patient_ID)-n_distinct(ctDNA1$Patient_ID)` patients are eliminated because of this step. There are  `r no.patient.sucess` patients in the data with Sample Status being Success. 